---
title: "lab3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing the Data
```{r}
library(readr)
tents <- read_csv("tents.csv", col_types = "ccnniccc")
```

### Questions
1. Run `str()` on `tents`.  What information does it tell you? This is a useful function to know.

str(tents)

The str function seems to provide a summary of the structure of the given data set. For readr, it displays the columns (8), the data types of these columns, and a shortened list of observations (90) under each column.

2. What class is `tents`?

class(tents)

3. How many rows are in `tents`?

nrow(tents)

4. How many columns are in `tents`?

ncol(tents)

5. What are the names of the columns?

colnames(tents)

6. Display the first 5 rows.

head(tents,5)

7. Display the last 4 rows.

tail(tents,4)

8. Find out how to use `quantile` to find the interquartile range (75th percentile - 25th percentile) of the weights of the tents.

quantile(tents$weight) 

9. Find the standard deviation of the heights of the `tents`.  Does `R` use `n` or `n-1` in the denominator when calculating standard deviation?

sd(tents$height)

R uses n-1 in the denominator when calculating the standard deviation.

10. The weight is given in grams.  Add a new column to the data frame called `weight_lbs` that contains the weight of the tents in pounds.  Remember to overwrite `tents` with your update.

tents$weight_lbs <- tents$weight/453.6

tents$weight_lbs

11. The height is given in centimeters.  Add a new variable `height_in` to the data frame that contains the height of the tents in inches. Remember to overwrite `tents` with your update.

tents$height_in <- tents$height/2.54

tents$height_in

12. How many tents cost more than $400?

length(which(tents$price>400))

13. Subset the data frame so that it only contains tents that are valued $300 or less.  Assign the subsetted data frame the name `so_cheap_tents`.

so_cheap_tents <- subset(tents, tents$price<=300)


14. Of the tents that are more than $300, what percent are `big-agnes` tents?

tents_over_300 <- subset(tents, tents$price>300)

tents_over_300

which (tents_over_300$brand == "big-agnes")

length (which (tents_over_300$brand == "big-agnes"))/nrow(tents_over_300)

15. Select the `bestuse` column of `tents` in four different ways: using `$`, `[[]]`, position number and `[]`, and variable name and `[]`. 

tents$bestuse

tents[["bestuse"]]

tents[[6]]

tents["bestuse"]

16. What is the name of the tent with the best height-to-weight ratio?

sort(tents$height/tents$weight, decreasing = TRUE)

17. Run `table` with two arguments, `seasons` and `capacity` (this should be one table).  What did it do?

table(tents$seasons, tents$capacity)

The command created a table with categories of seasons as the rows and categories of capacity as the columns.The cell entries were counts of tents that adhered to the stipulations of the rows and columns they were in.

18. What type of tent (bestuse) is the most common? Your answer should return only the use and not the count.

Backpacking was the most common best use. 

table(tents$bestuse)

19. Create a data frame called `kelty` with only data on kelty tents.  Do the same for `big-agnes`.  Then, combine the two data frames.

kelty <- subset(tents, tents$brand == "kelty")

big_agnes <- subset(tents, tents$brand == "big-agnes")

kelty_and_big_agnes <- rbind(kelty, big_agnes)

kelty_and_big_agnes
