---
title: "lab9"
author: "Andrew Do"
date: "June 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

The following libraries are allowed for this lab.  They have already been loaded in this document.

* `stringr`
* `dplyr`
* `tidyr`

```{r, include = FALSE}
library(stringr)
library(dplyr)
library(tidyr)
```


## Our Glorious Foundation

The vector pioneers has already been created for you
```{r}
pioneers <- c("GAUSS:1777", "BAYES:1702", "NEYMAN:1894", "LAPLACE:1749")
```

1.  Review: Split the names from the birth years using `str_split`, discarding the colon.  Assign the output to the variable `split_pioneers`.
```{r}
# Your code
split_pioneers <- str_split(pioneers, ":")
split_pioneers
```

2.  What class is the output object?
```{r}
# Your code
class(split_pioneers)
```

3.  Without writing a loop, convert the strings to first-letter uppercase only.  Hint: `str_to_upper` is actually part of a family of functions.  Getting the documentation on it will reveal other functions like it.  Assign the output to the variable `statisticians`
```{r}
# Your code
statisticians <- lapply(split_pioneers, str_to_title)
statisticians
```

4.  Try running `rbind` on statisticians.  What happens?
```{r}
# Your code
rbind(statisticians)
```
It creates a bizarre 1x4 matrix with vector elements.

5.  Using `do.call`, construct an `rbind` call that will turn the list into a matrix, `pioneers_df`
```{r}
# Your code
pioneers_df <- do.call (rbind, statisticians)
pioneers_df

```

6.  Review: Convert the matrix into a data frame and name the columns `name` and `byear`.  Hint: `as...`
```{r}
# Your code
df <- data.frame(pioneers_df, stringsAsFactors=F)
colnames(df) <- c("name", "byear")
df
```

7.  Review: Convert the column types of `name` and `byear` to character and integer, respectively.  Hint: Subset and `as...` 
```{r}
# Your code
df[, "name"] <- as.character(df[, "name"])
df[, "byear"] <- as.integer(df[, "byear"])
```

## Narcotics

The `sleep` dataset (built into R---no need to load anything) shows the effect of soporific (increases in sleep) drugs as compared to a control group.  There are two drugs represented in the data, as denoted by the group number.  The ID number is unique to each patient.

1.  Use `tapply` to find the average change in sleep hours for each of the drugs.
```{r}
# Your code
tapply(sleep$extra, sleep$group, mean)
```

2.  Use `lapply` and an anonymous function to split the data frame into 10 2x3 sub-data frames grouped by ID.  Assign the output to `split_sleep`.  This is just practice thinking about `lapply` since you could actually just use this call in practice:
```{r}
split(sleep, sleep$ID)

```

Hint: The first argument of `lapply` can be `1:10`.
```{r}
# Your code
split_sleep <- lapply(1:10, function(x) sleep[sleep$ID == x,])
```

3.  Use `vapply` on `split_sleep` to calculate the difference in extra sleep (group 2 - group 1) for each patient.  Hint: You can be a little fancy here and use `with` as your `FUN`.  Remember to give `vapply` a suitable template vector.
```{r}
#vapply(split_sleep, with (sleep[sleep$ID == 1:10], (split_sleep$group == 2] - split_sleep$group == 1]))
#q
```

## Look Me in the Iris

Iris is a rather famous data set introduced by Ronald Fisher in his paper _The use of multiple measurements in taxonomic problems_ as an example of classification by dividing the data using (hyper)-planes.  It is already pre-loaded into `R` under `iris` (no package needed) and is often used as an example dataset in documentation.  I suggest inspecting the data set before continuing.

Use the data wrangling verbs of `dplyr` in conjunction with the piping operator `%>%` to do the following tasks.  Your output should be data frames assigned to the names in bold.

1.  Get only the sepal and petal length columns of the data frame.  __iris_len__
```{r}
# Your code
select(iris, Sepal.Length, Petal.Length) -> `__iris_len__`
`__iris_len__`

```
2.  Get every column of the data frame besides Petal length and width.  Do not call on the other three columns by their names.  __no_petal__
```{r}
# Your code
select(iris, -Petal.Length, -Petal.Width) -> `__no_petal__`
`__no_petal__`

```
3.  Get only the species column, renaming it `species_name`.  __species__
```{r}
# Your code
species <- select(iris, Species)
rename(species, species_name = Species) -> `__species__`
`__species__`
```
4.  Create a column called `length_sum` that contains the sum of the petal and sepal lengths.  __add_length__
```{r}
# Your code
mutate(iris, length_sum = Petal.Length + Sepal.Length) -> `__species__`
`__species__`
```
5.  Get only the rows of the data frame where the species is `setosa`. __only_setosa__
```{r}
# Your code
filter(iris, Species == "setosa") -> `__only_setosa__`
`__only_setosa__`
```
6.  Get only the rows of the data frame where the species name contains exactly 4 vowels and the sum of the sepal length and width is less than 8.  __nonsense_filter__
```{r}
#species_name <- rownames(iris$Species)
#four_vowels <- "(.*[aeiou]){4}"
#species_with_four_vowels <- str_subset(species_name, four_vowels)
#filter(species_with_four_vowels, Sepal.Length + Sepal.Width <8)
#q
```
7.  Get the average sepal width `avg_sw` and average sepal length `avg_sl` for each of the 3 species.  __sepal_avg_by_species__
```{r}
#iris  %>%
  #different_species <- group_by(Species)  %>%
  #sepal_avg_by_species <- summarise(average = mean(Sepal.Width, Sepal.Length))
#sepal_avg_by_species
#q
```
8.  Overwrite the `Sepal.Width` column so that it reads `"wide"` if the width is greater than 3 and `"narrow"` otherwise.  Do the same for `Sepal.Length` except the threshold is 5.5 and the replacement is `"long"` and `"short"`.  The `ifelse` command might be useful.  __factorized__
```{r}
# Your code
#greater_than_3 <- iris$Sepal.Width > 3
#3_or_less <- iris$Sepal.Width <= 3
#rename(iris, wide = greater_than_3, narrow = 3_or_less)

#greater_than_5.5 <- iris$Sepal.Width > 5.5
#5.5_or_less <- iris$Sepal.Width <= 3
#rename(iris, long = greater_than_5.5, short = 5.5_or_less)
#q
```
9.  Count how many of each wide-long, wide-short, narrow-long, and narrow-short pairs there are.  Hint: `group_by` followed by `tally` or `summarise`.  __sepal_counts__
```{r}
# Your code
#pairs <- group_by(Sepal.Width, Sepal.Length)
#tally(pairs)
#q
```

