---
title: "lab1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data

```{r raiders}
scored <- c(14, 14, 9, 14, 28, 13, 13, 24, 17, 6, 24, 0, 24, 13, 26, 14)
against <- c(19, 30, 16, 38, 31, 24, 23, 30, 41, 13, 20, 52, 13, 31, 24, 47)
```

# Part 1

1. How many points did the Raiders score in game 7?

scored[7]

2. What was the score of the first 5 games?

head(scored,5)

3. What were the scores of the even games? Try to do this problem without typing in c(2,4,...)

scored[seq(2, length(scored), 2)]

4. What was the score of the last game? Try to do this problem without typing in the number of the last game.

tail(scored,1)

5. Select the scores higher than 14.

which(scored>14)

6. Select the scores equal to 14.

which(scored==14)


7. Select the scores equal to 13 or 14.

which(scored==13|scored==14)

8. Select the scores between between 7 and 21 points.

which(scored >=7 & scored <=21)

9. Sort the scores in decreasing order.

sort(scored, decreasing=TRUE)

10. What was the highest score achieved?

max(scored)

11. What was the lowest score achieved?

min(scored)

12. What was the average score?

mean(scored)

13. Use the `summary()` function to give some summary statistics.

summary(scored)

14. In how many games did the Raiders score over 20 points?

which(scored>=20)

scoredover20<-scored[c(5, 8, 11, 13, 15)]

length(scoredover20)

# Part 2
These questions refer to both vectors.

1. What was the sum of the scores in game 7?

scored[7]+against[7]

2. What was the sum of the scores in each of the last 5 games?

tail(scored,5)+tail(against,5)

3. What was the difference in scores in each of the odd games?

scored[seq(1,length(scored),2)]-against[seq(1,length(against),2)]

4. Which games did the Raiders win?

which(scored-against>=0)

5. Were there any ties?

which(scored-against==0)