---
title: "homework1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install the `nycflights13` package by typing `install.packages("nycflights13")` into the console (do not leave this in your R Markdown document).  

```{r}
library(nycflights13)
data(flights)
```

1. How many rows are in this data frame? __num_rows__

num_rows <- nrow(flights)

num_rows

2. How many variables are in this data frame? __num_vars__

num_vars <- ncol(flights)

num_vars

3. What was the longest departure delay? __longest_dep_delay__

longest_dep_delay <- which.max(flights$dep_delay)

longest_dep_delay

4. How long were the top 10 longest flights measured using `air_time`? __longest_flights__

ranking_flight_air_times <- sort(flights$air_time, decreasing = TRUE)

longest_flights <- head(ranking_flight_air_times,10)

longest_flights˚

5. What is the average departure delay time for United Airlines (carrier UA)? __ua_delay__

Received NA as the result for the average due to the presence of "NA" values in the data for departure delay times for United.

UA_flights <- subset(flights, flights$carrier == "UA")

ua_delay <- mean(UA_flights$dep_delay)

ua_delay

6. Order the months from highest average departure delay time to lowest.  __month_delay__

My intention for this question was to create subsets of the data for each month, isolate the column for departure delay times to compute an average, and order the monthly averages by manually sorting (copying and pasting the obtained results). However, based on my progress so far, it seems that the mean will return "NA" for all of the monthly averages because of the presence of "NA" in the data when used to compute the mean.  

January_flights <- subset(flights, flights$month == "1")

mean(January_flights$dep_delay)

February_flights <- subset(flights, flights$month == "2")

mean(February_flights$dep_delay)

March_flights <- subset(flights, flights$month == "3")

mean(March_flights$dep_delay)

April_flights <- subset(flights, flights$month == "4")

mean(April_flights$dep_delay)

May_flights <- subset(flights, flights$month == "5")

mean(May_flights$dep_delay)

June_flights <- subset(flights, flights$month == "6")

mean(June_flights$dep_delay)

July_flights <- subset(flights, flights$month == "7")

mean(July_flights$dep_delay)

August_flights <- subset(flights, flights$month == "8")

mean(August_flights$dep_delay)

September_flights <- subset(flights, flights$month == "9")

mean(September_flights$dep_delay)

October_flights <- subset(flights, flights$month == "10")

mean(October_flights$dep_delay)

November_flights <- subset(flights, flights$month == "11")

mean(November_flights$dep_delay)

December_flights <- subset(flights, flights$month == "12")

mean(December_flights$dep_delay)

monthly_averages <- c(mean(January_flights$dep_delay), mean(February_flights$dep_delay), mean(March_flights$dep_delay), mean(April_flights$dep_delay), mean(May_flights$dep_delay), mean(June_flights$dep_delay), mean(July_flights$dep_delay), mean(August_flights$dep_delay), mean(September_flights$dep_delay), mean(October_flights$dep_delay), mean(November_flights$dep_delay), mean(December_flights$dep_delay))

month_delay <- sort (monthly_averages, decreasing = TRUE)

month_delay

7. How many unique planes does American Airlines (AA) have? __num_planes_aa__

American_airlines <- subset(flights, flights$carrier == "AA")

num_planes_aa <- length(unique(American_airlines$tailnum))

num_planes_aa

8. How many flights originating from JFK flew more than 1000 miles? __jfk_1000mi__

JFK_flights <- subset (flights, flights$origin == "JFK")

jfk_1000 <- length(JFK_flights$distance > 1000)

jfk_1000

9. How many airlines fly from LGA to DFW? __num_lga_dfw__

LGA_flights <- subset (flights, flights$origin == "LGA")

num_lga_dfw <- length (LGA_flights$dest == "DFW")

num_lga_dfw

10. Which airline had the most flights coming out of JFK? __most_flights_jfk__

The airline corresponding with 42076 flights, B6, had the most flights coming out of JFK. 

Flights_jfk <- subset(flights, flights$origin == "JFK")

length(which(Flights_jfk$carrier == "AA"))

length(which(Flights_jfk$carrier == "9E"))

length(which(Flights_jfk$carrier == "B6"))

length(which(Flights_jfk$carrier == "DL"))

length(which(Flights_jfk$carrier == "EV"))

length(which(Flights_jfk$carrier == "HA"))

length(which(Flights_jfk$carrier == "MQ"))

length(which(Flights_jfk$carrier == "UA"))

length(which(Flights_jfk$carrier == "US"))

length(which(Flights_jfk$carrier == "VX"))

most_flights_jfk <- max(length(which(Flights_jfk$carrier == "AA")), length(which(Flights_jfk$carrier == "9E")), length(which(Flights_jfk$carrier == "B6")), length(which(Flights_jfk$carrier == "DL")), length(which(Flights_jfk$carrier == "EV")), length(which(Flights_jfk$carrier == "HA")), length(which(Flights_jfk$carrier == "MQ")), length(which(Flights_jfk$carrier == "UA")), length(which(Flights_jfk$carrier == "US")), length(which(Flights_jfk$carrier == "VX")))

most_flights_jfk